---
title: Intensity and response times
author: "Bruno Nicenboim"
date: "`r Sys.Date()`"
slug: custom-lik
categories: ["stats"]
tags: ["stan", "Bayesian", "r"]
draft: true 
---

First part d*, p*, *q*, r* like functions for R. MLE
Second part brms
Third part custom funciton Stan and SBC
Fourth part C++ implementation

```{r}
ddm <- function(x, delta, epsilon, t_0){
  ((delta + 1) * (epsilon - 1)/
     ((delta + epsilon) * t_0)) *
    ifelse(x <= t_0,
    (x/ t_0)^delta,
    (x / t_0)^(-epsilon))
}
```
In lieu of  any suitable  theoretical al-
ternative, we  attempted  to  fit  the  doublemonomial (DM) to our  RT  distributions.The  DM  was  derived  by Luce  and  Galan-ter (1963,  p.  288) as  the  form  of  a  gen-
eralization  function  in  magnitude estima-
tion. It  consists  of  two  power  functions
that  meet  at  the  mode. As a  result  it  hastwo  qualitative  features  of  our  empiricaldistributions—peakedness  at  the  mode and
high tails (in the  sense  that  a negativepower  approaches  zero much  more  slowlythan  an  exponential). The  density func


https://www.wolframalpha.com/input/?i=integrate+%5B%28%28delta+%2B+1%29+*+%28epsilon+-+1%29%2F++++++%28%28delta+%2B+epsilon%29+*+t_0%29%29+++++%28x%2F+t_0%29%5Edelta%5D+

(x (ϵ - 1) (x/t_0)^δ)/(t_0 (δ + ϵ)) + constant

https://www.wolframalpha.com/input/?i=integrate+%5B%28%28delta+%2B+1%29+*+%28epsilon+-+1%29%2F++++++%28%28delta+%2B+epsilon%29+*+t_0%29%29+++++%28x%2F+t_0%29%5E%28-epsilon%29%5D+from+x+%3D+t_0+to+x

integral_(t_0)^x (((δ + 1) (ϵ - 1)) (x/t_0)^(-ϵ))/((δ + ϵ) t_0) dx = ((δ + 1) x^(-ϵ) (t_0 x^ϵ - x t_0^ϵ))/(t_0 (δ + ϵ)) for t_0>0 ∧ Re(x)>t_0 ∧ Im(x) = 0

```{r}
pdm <- function(x, delta, epsilon, t_0){

lower_than_t_0 <- function(x) (x* (epsilon - 1) * (x/t_0)^delta)/(t_0 *(delta + epsilon))

  ifelse(x <= t_0, lower_than_t_0(x),
         lower_than_t_0(t_0) + 
  ((delta + 1)* x^(-epsilon) *(t_0* x^epsilon - x * t_0^epsilon))/(t_0 *(delta + epsilon)) 
    )
}
pdm(10000000000000000000000000000000, 1.1,1.1,200)
```



number generation:

https://mc-stan.org/docs/2_27/stan-users-guide/truncated-random-number-generation.html


