---
title: "CURRICULUM VITAE"
output: 
  pdf_document:
    toc: false # table of content true
    number_sections: true  ## if you want number sections at each table header
    keep_tex: true	
---


# Contact Information
 
```{r setup, include=FALSE}
# remove the year from talks, add months, order by months
# 
library(knitr)
library(rmarkdown)
library(htmltools)
source("common.R")

options(htmltools.dir.version = FALSE)

library(RefManageR)
bib <- ReadBib("mypub.bib", check = F)
 ## 'style' should be one of "text", "Bibtex", "citation", "html", "latex", "textVersion", "R", "Biblatex", "markdown", "yaml"
BibOptions(check.entries = FALSE, style = "html", cite.style = "authortitle",
           bib.style = "authoryear",sorting="ydnt",max.names =10,dashed=FALSE, no.print.fields=c("issn","url","customa","customb","customc"))


sort(bib,sorting="ydnt")
endcit<- "</cite></p>"

this_year <- as.numeric(format(Sys.Date(), "%Y"))
first_year <-  2000

```
--------------        -------------------------------------------
Email:                bruno.nicenboim@uni-potsdam.de
Tel:                  (+49)-0331/977-2796
Office:               House 14, Room 214, Campus Golm, Potsdam
Address:              Universität Potsdam
                      Department für Linguistik
                      Haus 14
                      Karl-Liebknecht-Straße 24-25
                      D-14476 Potsdam
--------------        -------------------------------------------


<br/>
<br/>



# Education

------------------------  -----------------------------------------------------------------------
2011--2017                **PhD  in Cognitive Science, University of Potsdam**

                          Topic: Dependency resolution as a retrieval process

                          Supervisors: Prof. Dr. Shravan Vasishth and Prof. Dr. Reinhold Kliegl

                          Grade: *summa cum laude* <br/> <br/> 


2010                      M.A. in Linguistics, Tel Aviv University

                          Topic: Processing Complex NP Islands in Hebrew

                          Supervisors: Prof. Dr. Julia Horvath and Prof. Dr. Tal Siloni 

                          Grade: *summa cum laude*  <br/> <br/> 


2009                      B.A Double Major, Tel Aviv University

                          Linguistics. Grade: *summa cum laude*

                          Sociology and Anthropology. Grade: very good
------------------------  -----------------------------------------------------------------------

<br/>
<br/>

# Academic-related work 

------------------------  -----------------------------------------------------------------------
2017--                    **Postdoctoral Researcher**, University of Potsdam.

                          PIs: Shravan Vasishth, Frank Rösler <br/> <br/> 


2016--2017                Coordinator of the Master in Cognitive Systems,  University of Potsdam <br/><br/>  
    


2009--2009                Research assistant, Tel Aviv University. Syntax and Lexicon interface. 

                          PIs: Prof. Dr. Tal Siloni and Prof. Dr. Julia Horvath <br/> <br/> 


2008--2009                Research assistant, Tel Aviv University. Psycholinguistics.  

                          Assistance in experiments using E-prime and CHILDES software. 

                          PI: Prof. Dr. Ruth Berman  
---------------------------------   --------------------------------------------------------------------------------


<br/> 
<br/>

# Articles

## Published <!-- or accepted --> in peer-reviewed journals

```{r articles-cv,echo=F,results="asis",cache=F}
articlesall <- bib[list(bibtype="article")]
last_year <- max(as.numeric(articlesall$year),na.rm=T)
library(purrr)
for(y in seq(last_year,first_year,-1)){
    articles <- suppressMessages(articlesall[list(year=as.character(y))])
    if(length(articles)!=0){    
        ## cat(paste0(" \n" )) 
        walk(articles, ~ {
            cat("* ")
            output(.x,withbrackets=F)
            cat("\n")
        }) 
    }
}    
 
```
<br/>
<br/>


## Unpublished manuscripts 

```{r,echo=F,results="asis",cache=F}
 
articles <- bib[list(bibtype="unpublished")]

walk(articles, ~ {
    cat("* ")
    output(.x,withbrackets=F)
    cat("\n")
}) 

```

<br/>
<br/>

# Conferences

```{r,echo=F,results="asis",cache=F}
presentationsall <- bib[list(bibtype="inproceedings")]
 last_year <- max(as.numeric(presentationsall$year))

for(y in seq(last_year,first_year,-1)){

    presentations <- suppressMessages(presentationsall[list(year=as.character(y))])
    if(length(presentations)!=0){    
        ## cat(paste0("## ",as.character(y),"\n" ))  
        ## cat(paste0(" \n" )) 
        walk(presentations, ~ {
            cat("* ")
            output(.x,withbrackets=F)
            cat("\n")
        }) 
       ## output(presentations,withbrackets=F)
    }
}
```
<br/>
<br/>

# Invited talks

```{r,echo=F,results="asis",cache=F}
 
library(stringr)

talksall <- bib[list(bibtype="misc")]

for(y in seq(this_year,first_year,-1)){

    talks <- suppressMessages(talksall[list(year=as.character(y))])
    if(length(talks)!=0){    
        ## cat(paste0("## ",as.character(y),"\n" ))  
        ## output(talks,withbrackets=F,hack=T)
        walk(talks, ~ {
            cat("* ")
            output(.x,withbrackets=F,hack=T)
            cat("\n")
        }) 
    }
}

```
<br/>
<br/>


# Honors and grants

------------------------  -----------------------------------------------------------------------
2012--2018                Potsdam Graduate School (PoGS) Travel Grants 

2014                      Kommission für Forschung und wissenschaftlichen Nachwuchs (FNK, University of Potsdam) Travel Grant

2014                      Potsdam Graduate School (PoGS) PhD Completion Scholarship (5 months)

2013                      Minerva Fellowship Extension for Doctoral Research (1 year)

2011                      Minerva Fellowship for Doctoral Research (2 years)

2009                      Posis Scholarship from the School of Cultural Studies, Tel Aviv University

2008                      Faculty of Humanities Scholarship for Achievements in M.A. studies, Tel Aviv University

2006                      Dean's Award for Achievements in Linguistics, Tel Aviv University 
------------------------  -----------------------------------------------------------------------

<br/>
<br/>

# Teaching experience

## Workshops and summer schools

------------------------  -----------------------------------------------------------------------
2020 (Forthcoming)        "Introduction to computational Bayesian methods using Stan", 
                             with Shravan Vasishth 
                       
                          Physalia courses, Berlin, Germany\vspace{.1cm}
                          
2020 (Forthcoming)        "Methods in Advanced Statistics", with Shravan Vasishth 
                       
                          2020 Winter School 
                          
                          by Netherlands Graduate School in Linguistics (LOT),
                          
                          Tilburg, Netherlands\vspace{.1cm}

2019 (Forthcoming)        "Introduction to Bayesian statistics using brms", 

                          University of Cologne, Germany\vspace{.1cm}

2019 (Forthcoming)        "Introduction to Bayesian statistics using brms", 

                          University of Edinburgh, UK\vspace{.1cm}
                          
2019 (Forthcoming)        "Advanced Bayesian methods",

                          Third Summer School on Statistical Methods for Linguistics and Psychology,

                          University of Potsdam, Germany\vspace{.1cm}
                          
2018                      Talk:"Cognitive models of memory processes in sentence comprehension:  A case study using Bayesian hierarchical modeling"  

                          Masterclass in Bayesian Statistics, Research school,

                          CIRM (Marseille Luminy, France)\vspace{.1cm} 


2018                      "Advanced topics in Bayesian modeling",

                          Second Summer School on Statistical Methods for Linguistics and Psychology,

                          University of Potsdam, Germany\vspace{.1cm} 


2017                      "Introduction to Bayesian Modeling using Stan", 

                          13. Tagung der Fachgruppe Methoden und Evaluation der Deutschen Gesellschaft für Psychologie,

                          Tübingen, Germany\vspace{.1cm} 

2017                      "Introduction to Bayesian modeling",

                          First Summer School on Statistical Methods for Linguistics and Psychology,

                          University of Potsdam, Germany\vspace{.1cm} 
------------------------  -----------------------------------------------------------------------


## Courses

------------------------  -----------------------------------------------------------------------
2015--2017 (Winter)       **Lecturer** in "Advanced Data Analysis". University of Potsdam

2016 (Summer)             **Lecturer** in "Predictions in Language Processing".  University of Potsdam

2015--2016 (Winter)       **Lecturer** in "Individual Differences in Sentence Processing".  University of Potsdam

2015 (Summer)             **Lecturer** in "Predictions in Language Processing".  University of Potsdam

2008--2010                **Teaching assistant** in "Syntax Beginners" and "Foundations of Theoretical Linguistics" Tel Aviv University
------------------------  -----------------------------------------------------------------------



<br/>
<br/>


# Supervising

------------------------  -----------------------------------------------------------------------
2019			          Chiara Tschirner (co-supervision with Prof. Dr. Vasishth), Master Thesis.	

2018                      Eva-Maria Fey (co-supervision with Prof. Dr. Vasishth), Bachelor Thesis. Topic: "The effect of individual differences and the functional organisation of the human brain in joke comprehension based on left- and right-handedness"


2016                      Daniel Grünke (co-supervision with Prof. Dr. Vasishth), Bachelor Thesis. Topic: "Der Zusammenhang zwischen Arbeitsgedächtnis und Satzverarbeitung" 
 

2015                      Lisa Münchberger (co-supervision with Prof. Dr. Vasishth), Bachelor Thesis. Topic: "Unflüssigkeiten und Hemmungen in der Sprachverarbeitung im Deutschen" 
------------------------  -----------------------------------------------------------------------


<br/>
<br/>


# Open software 

* [eeguana](https://bnicenboim.github.io/eeguana/). An R package for flexible manipulation of EEG data. (https://bnicenboim.github.io/eeguana/)


<br/>
<br/>


# Academic-related training

------------------------  -----------------------------------------------------------------------
2016                      Groningen Spring School on Cognitive Modeling, the Netherlands

2015--2016                International Teaching Professionals,  University of Potsdam, Germany 

2013                      Autumn School Methods in Language Comprehension,  Rovereto (TN), Italy 

2012                      Eye-tracking-while-reading  at Laboratory of Integrative Neuroscience, University of Buenos Aires (3 months). Host: Dr. Mariano Sigman 
------------------------  -----------------------------------------------------------------------





<br/>
<br/>




# Reviewing


* Journal of Experimental Psychology: Learning, Memory, and Cognition

* Neuropsychologia

* Lingua

* Phonological Data and Analysis

* Language, Cognition and Neuroscience

* Routledge Manuscript

* Journal of Memory and Language

* Quarterly Journal of Experimental Psychology

* Journal of Cognitive Psychology  

* Journal of Phonetics 

* CUNY Conference

* Quantitative approaches in corpus linguistics and psycholinguistics, Paris, France    

<br/>
<br/>

# Languages

* Spanish (Mother tongue)

* English   (Fluent)

* Hebrew    (Fluent)

* German    (Advanced)
 


<br/>
<br/>

# Non-academic professional experience

------------------------  -----------------------------------------------------------------------
2009--2011                QA engineer: NLP Testing at Clearforest.                           Testing rules and heuristics for identifying semantic entities and relations in English, Spanish and French texts. 

2007--2009                Content Specialist position at Celebros. Organization, analysis and categorization of data of Spanish and U.S. clients' databases. 
------------------------  -----------------------------------------------------------------------



